The project for the statistical inference class
========================================================
The project consists of two parts:

1. Simulation exercises.
2. Basic inferential data analysis.

# PART 2. Basic inferential data analysis.

## Load the ToothGrowth data and perform some basic exploratory data analyses

``` {r data}
library(datasets)
data = ToothGrowth
str(data)

library(ggplot2)

qplot(dose, len, data=data, color = supp, size = I(3), alpha = I(.7), main = "Len by Dose") +
    geom_smooth(method=lm, se = FALSE)

qplot(dose, len, data = data, facets = .~supp, size = I(3), color = supp, main = "Len by Dose for each Supp")

qplot(supp, len, data = data, color = dose, size = I(3), main = "Len by Supp")

qplot(supp, len, data = data, color = dose, size = I(3), facets = .~dose, main = "Len by Supp for each Dose")
```

## Provide a basic summary of the data.
Summarizing data by supp and dose.  
``` {r datasummary}
# By supp
OJ <- data[data$supp == "OJ", "len"]
oj.mean <- mean(OJ)
oj.sd <- sd(OJ)
VC <- data[data$supp == "VC", "len"]
vc.mean <- mean(VC)
vc.sd <- sd(VC)

# By dose
d05 <- data[data$dose == 0.5, "len"]
d05.mean <- mean(d05)
d05.sd <- sd(d05)
d1 <- data[data$dose == 1, "len"]
d1.mean <- mean(d1)
d1.sd <- sd(d1)
d2 <- data[data$dose == 2, "len"]
d2.mean <- mean(d2)
d2.sd <- sd(d2)

g1 <- qplot(OJ, binwidth = 0.5)
g2 <- qplot(VC, binwidth = 0.5)
g3 <- qplot(d05, binwidth = 0.5)
g4 <- qplot(d1, binwidth = 0.5)
g5 <- qplot(d2, binwidth = 0.5)

library(gridExtra)
grid.arrange(g1, g2, g3, g4, g5, ncol = 2)
```
  
VC: mean `r vc.mean`, sd `r vc.sd`.  
OJ: mean `r oj.mean`, sd `r oj.sd`.  
d05: mean `r d05.mean`, sd `r d05.sd`.  
d1: mean `r d1.mean`, sd `r d1.sd`.  
d2: mean `r d2.mean`, sd `r d2.sd`.  

## Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose.