---
output: html_document
---
The project for the Statistical Inference class
========================================================
*GitHub Repo (if source code required): https://github.com/d2ski/sicp1*

## PART 1. Simulation.

The exponential distribution simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also also 1/lambda. Lambda = 0.2 for all of the simulations. The distribution of averages of 40 exponential(0.2)s. A thousand simulated averages of 40 exponentials (n).

### Show where the distribution is centered at and compare it to the theoretical center of the distribution.

Simulating the distribution of averages of 40 exponential(0.2)s:  

```{r data, echo=FALSE}
# Data frame of averages from 1000 simulations
nosim <- 1000
lambda <- 0.2
nsamp <- 40
means <- c()
sds <- c()
set.seed(177)

for (n in 1:nosim) {
    sample <- rexp(nsamp, lambda)
    means <- c(means, mean(sample))
    sds <- c(sds, sd(sample))
}
data <- data.frame(1:nosim, means, sds)
names(data) <- c("nsim", "means", "sds")

# Histogram of the distribution with its mean
library(ggplot2)

mdistr <- mean(data$means)

g <- ggplot(data, aes(x=data$means)) + geom_histogram(binwidth=.5, color="black", fill="blue") +
    geom_vline(aes(xintercept=mean(data$means)), color="red", linetype="dashed", size=1)
g
```

The center of the sample distribution is mean `r mdistr` (red dashed line).  
``` {r mean,echo=FALSE,results='hide'}
mteor <- lambda ^ (-1)
```
The theoretical center of the distribution (mean) is 1/lambda: `r mteor`.  
The mean of the sample distribution approximately equals to the theoretical mean.

### Show how variable it is and compare it to the theoretical variance of the distribution.

``` {r variance, echo=FALSE}
# Variance of the sample distribution
vardistr <- var(data$means)
# and theoretical variance
varteor <- (lambda ^ (-2)) / nsamp
```

The variance of the sample distribution is `r vardistr`. The theoretical variance is `r varteor`, which is calculated as varicance of exponential distribution divided by n. The variances are approximately the same.

### Show that the distribution is approximately normal.

``` {r normal, echo=FALSE, fig.height=3,fig.width=3,warning=FALSE}
g2 <- ggplot(data, aes(x=data$means)) + geom_histogram(aes(y=..density..), binwidth=.5,
                                                       color="black", fill="blue") +
    geom_vline(aes(xintercept=mean(data$means)), color="red", linetype="dashed", size=1) +
    geom_density(alpha=.2, fill="#FF6666", size=1.5)
g2

qqnorm(data$means, cex=0.3)
qqline(data$means)
```

### Evaluate the coverage of the confidence interval for 1/lambda.
I interpreted this question to mean: discover proportion of time when 95% confidence interval
contains the point estimate (the theoretical mean = 1/lambda). The plot below shows
when the point estimate is within 95% confidence level [`ll`, `ul`], where `ll` - lower limit,
`ul' - upper limit of the interval.

``` {r coverage, echo=FALSE,fig.height=4, fig.width=6}
# Calculating confidence interval for each simulation
ul <- c()
ll <- c()
for (n in data$nsim) {
    m <- data[n, "means"]
    s <- data[n, "sds"]
    ul <- c(ul, m + 1.96 * s / sqrt(nsamp))
    ll <- c(ll, m - 1.96 * s / sqrt(nsamp))
}

coverage <- round(sum((ll < mteor) & (ul > mteor)) / nosim, 2) * 100

# Plot
g3 <- ggplot(data.frame(ul, ll), aes(x=ll, y=ul, color=(ll < mteor) & (ul > mteor))) + geom_point(shape=1)
g3 <- g3 + geom_hline(aes(yintercept=mteor)) + geom_vline(aes(xintercept=mteor))
g3
```
  
The theoretical mean (1/lambda) = `mteor` = `r mteor`. Percent of confidence intervals containing the theoretical mean is `r coverage` %.